---
title: "Vowel Acoustics as Predictors of Speech Intelligibility in Dysarthria: Alternate Analysis"
output: html_notebook
---

This is the code for the statistical analysis for "Vowel Acoustics as Predictors of Speech Intelligibility in Dysarthria."

# Loading Packages

```{r}

library(rio)
library(tidyverse)
library(irr) # install.packages('irr')
library(performance)
library(car)
library(ggpubr)
library("Hmisc") # install.packages('Hmisc')
library(ggridges)
library(furniture) # install.packages('furniture')
library(gt)
library(patchwork)
library(ks)
library(emuR) # install.packages('emuR')
library(geomtextpath) # remotes::install_github("AllanCameron/geomtextpath")

```

# Upload Datasets

```{r}

Reliability <- rio::import("Prepped Data/Reliability Data.csv")
AcousticData <- rio::import("Prepped Data/AcousticMeasures.csv") %>%
  dplyr::mutate(intDiff = VAS - transAcc)

AcousticData <- AcousticData %>%
  dplyr::filter(!grepl("_rel", Speaker)) %>%
  dplyr::select(c(Speaker, Sex, Etiology, vowel_ED_b, VSA_b, Hull_b, Hull_bVSD_25, Hull_bVSD_50, Hull_bVSD_75, VAS, transAcc))

Listeners <- rio::import("Prepped Data/Listener_Demographics.csv") %>%
  dplyr::select(!c(StartDate:proloficID, Q2.4_6_TEXT, Q3.2_8_TEXT, AudioCheck:EP3))

Listeners$race[Listeners$Q3.3_7_TEXT == "Native American/ African amercing"] <- "Biracial or Multiracial"
```

# Alternate Analysis

In this alternate analysis, we are looking at the relationship between these acoustic measures with speech intelligibility for the ALS/PD and the HD/Ataxic speakers separately. We create a new variable called Incoord, where the ALS/PD Speakers are set as the reference group (in order to compare to the Ataxic/HD Speaker Group). Group Comparisons, additional data visualizations, and further linear model comparisons are completed.

## Data Prep

```{r}

# Creating a new variable in AcousticData. Incoord is a dummy variable. ALS/PD Speakers = 0, HD/Ataxic = 1

AcousticData <- AcousticData %>%
  dplyr::mutate(Incoord = case_when(Etiology == "HD" ~ 1,
                                    Etiology == "Ataxic" ~ 1,
                                    TRUE ~ 0)) %>%
  dplyr::mutate(Incoord = as.factor(Incoord))

```

## Descriptives 

```{r}

Descriptives_ALS.PD_incoord <- AcousticData %>%
  dplyr::group_by(Incoord) %>%
  dplyr::summarize(VSA_mean = mean(VSA_b, na.rm =T), VSA_sd = sd(VSA_b, na.rm = T),
                   Disp_mean = mean(vowel_ED_b, na.rm =T), Disp_sd = sd(vowel_ED_b, na.rm =T),
                   Hull_mean = mean(Hull_b, na.rm =T), Hull_sd = sd(Hull_b, na.rm =T),
                   VSD25_mean = mean(Hull_bVSD_25, na.rm =T), VSD25_sd = sd(Hull_bVSD_25, na.rm =T),
                   VSD50_mean = mean(Hull_bVSD_50, na.rm =T), VSD50_sd = sd(Hull_bVSD_50, na.rm =T),
                   VSD75_mean = mean(Hull_bVSD_75, na.rm =T), VSD75_sd = sd(Hull_bVSD_75, na.rm =T),
                   VAS_mean = mean(VAS, na.rm =T), VAS_sd = sd(VAS, na.rm =T),
                   OT_mean = mean(transAcc, na.rm =T), OT_sd = sd(transAcc, na.rm =T))

Descriptives_ALS.PD_incoord

```

## Data Vis

### Group Comparison Ridgeline Plots

```{r}

# VSA Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

VSA_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Area (Bark)") +
  theme_classic()

# Corner Dispersion Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

disp_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Corner Dispersion (Bark)") +
  theme_classic()

# Hull Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

hull_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Hull (Bark)") +
  theme_classic()

# VSD 25 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd25_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 25 (Bark)") +
  theme_classic()

# VSD 50 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd50_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_50,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 50 (Bark)") +
  theme_classic()

# VSD 75 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd75_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 75 (Bark)") +
  theme_classic()

# Orthographic Transcription Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

OT_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = transAcc,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Percent Words Correct") +
  theme_classic()

# Visual Analog Scale Intelligibility Rating Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

VAS_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Ratigns") +
  theme_classic()

# Creating Distributions by Incoord Figure

Distributions_incoord <- ggpubr::ggarrange(VSA_incoord.plot, disp_incoord.plot, hull_incoord.plot, vsd25_incoord.plot, vsd50_incoord.plot, vsd75_incoord.plot, OT_incoord.plot, VAS_incoord.plot,
                  ncol = 2,
                  nrow = 4)

# Saving Distribution Figure

ggsave("Plots/Distribution_incoord.png", plot = last_plot(), width = 10, height = 10, unit = "in")

```

### Scatterplots

```{r}

# OT Scatterplots by Incoord

OT_VSA.incoord <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Area (Bark)") +
  ylab("Percent Words Correct") +
  ggtitle("Orthographic Transcription") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

OT_disp.incoord <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Corner Dispersion (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_Hull.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Hull (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd25.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 25 (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd75.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 75 (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()


# VAS Scatterplots

VAS_VSA.incoord <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Area (Bark)") +
  ylab("VAS Score") +
  ggtitle("Visual Analog Scale") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

VAS_disp.incoord <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Corner Dispersion (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_Hull.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Hull (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd25.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 25 (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd75.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 75 (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()


# Creating Scatterplot Figure

Scatterplots_incoord <- ggpubr::ggarrange(OT_VSA.incoord, VAS_VSA.incoord, OT_disp.incoord, VAS_disp.incoord, OT_Hull.incoord, VAS_Hull.incoord, OT_vsd25.incoord, VAS_vsd25.incoord, OT_vsd75.incoord, VAS_vsd75.incoord,
                  ncol = 2,
                  nrow = 5)

## Saving Scatterplot Figure

ggsave("Plots/Scatterplots_incoord.png", plot = last_plot(), width = 15, height = 20, unit = "in", scale = .5)


```


## Group Comparisons

### Two Dataframes for Incoord Groups

```{r}

coord.group <- AcousticData %>%
  dplyr::filter(Incoord == 0)

incoord.group <- AcousticData %>%
  dplyr::filter(Incoord == 1)

```


### VSA

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(VSA_b[Incoord == 0]))
with(AcousticData, shapiro.test(VSA_b[Incoord == 1]))

## Equal Variance Check

res.ftest.VSA <- var.test(VSA_b ~ Incoord, data = AcousticData)
res.ftest.VSA

# Model Results

VSA_b_t <- t.test(incoord.group$VSA_b, coord.group$VSA_b, var.equal = T)
VSA_b_t

```

### Corner Dispersion

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(vowel_ED_b[Incoord == 0]))
with(AcousticData, shapiro.test(vowel_ED_b[Incoord == 1]))

## Equal Variance Check

res.ftest.disp <- var.test(vowel_ED_b ~ Incoord, data = AcousticData)
res.ftest.disp

# Model Results

disp_t <- t.test(incoord.group$vowel_ED_b, coord.group$vowel_ED_b, var.equal = T)
disp_t

```

### Hull

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_b[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_b[Incoord == 1]))

## Equal Variance Check

res.ftest.hull <- var.test(Hull_b ~ Incoord, data = AcousticData)
res.ftest.hull

# Model Results

hull_t <- t.test(incoord.group$Hull_b, coord.group$Hull_b, var.equal = T)
hull_t

```

### VSD 25

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_25[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_25[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd25 <- var.test(Hull_bVSD_25 ~ Incoord, data = AcousticData)
res.ftest.vsd25

# Model Results

vsd25_t <- t.test(incoord.group$Hull_bVSD_25, coord.group$Hull_bVSD_25, var.equal = T)
vsd25_t

```

### VSD 50

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_50[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_50[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd50 <- var.test(Hull_bVSD_50 ~ Incoord, data = AcousticData)
res.ftest.vsd50

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

vsd50_MW <- wilcox.test(Hull_bVSD_50 ~ Incoord, data = AcousticData)
vsd50_MW

```

### VSD 75

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_75[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_75[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd75 <- var.test(Hull_bVSD_75 ~ Incoord, data = AcousticData)
res.ftest.vsd75

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

vsd75_MW <- wilcox.test(Hull_bVSD_75 ~ Incoord, data = AcousticData)
vsd75_MW


```

### Orthographic Transcription Scores

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(transAcc[Incoord == 0]))
with(AcousticData, shapiro.test(transAcc[Incoord == 1]))

## Equal Variance Check

res.ftest.OT <- var.test(transAcc ~ Incoord, data = AcousticData)
res.ftest.OT

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

OT_MW <- wilcox.test(transAcc ~ Incoord, data = AcousticData)
OT_MW

```

### VAS

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(VAS[Incoord == 0]))
with(AcousticData, shapiro.test(VAS[Incoord == 1]))

## Equal Variance Check

res.ftest.VAS <- var.test(VAS ~ Incoord, data = AcousticData)
res.ftest.VAS

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

OT_MW <- wilcox.test(VAS ~ Incoord, data = AcousticData)
OT_MW

```

## OT Analysis

Since we found significant group differences for some acoustic measures between the ALS/PD and Ataxic/HD groups, we continued the heirarichal regression approach from OT Model 5. Adding in the Incoord predictor along with the interactions between the acoustic measures did not significantly improve model fit. So our original final OT model is retained.

### Model 6

```{r}

## Specifying Model 6

OT_Model6 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + Incoord, data = AcousticData)

## Model 6 Assumption Check

performance::check_model(OT_Model6)

## Model 6 Summary

summary(OT_Model6)

## Model 5 and Model 6 Comparison

anova(OT_Model5, OT_Model6)

```
### Model 7

```{r}

## Specifying Model 7

OT_Model7 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25, data = AcousticData)

## Model 7 Assumption Check

performance::check_model(OT_Model7)

## Model 7 Summary

summary(OT_Model7)

## Model 6 and Model 7 Comparison

anova(OT_Model6, OT_Model7)

```

### Model 8

```{r}

## Specifying Model 8

OT_Model8 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75, data = AcousticData)

## Model 8 Assumption Check

performance::check_model(OT_Model8)

## Model 8 Summary

summary(OT_Model8)

## Model 7 and Model 8 Comparison

anova(OT_Model7, OT_Model8)

```

### Model 9

```{r}

## Specifying Model 9

OT_Model9 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b, data = AcousticData)

## Model 9 Assumption Check

performance::check_model(OT_Model9)

## Model 9 Summary

summary(OT_Model9)

## Model 8 and Model 9 Comparison

anova(OT_Model8, OT_Model9)

```

### Model 10

```{r}

## Specifying Model 10

OT_Model10 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b, data = AcousticData)

## Model 10 Assumption Check

performance::check_model(OT_Model10)

## Model 10 Summary

summary(OT_Model10)

## Model 9 and Model 10 Comparison

anova(OT_Model9, OT_Model10)

```

### Model 11
mo
```{r}

## Specifying Model 11

OT_Model11 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b + Incoord*vowel_ED_b, data = AcousticData)

## Model 11 Assumption Check

performance::check_model(OT_Model11)

## Model 11 Summary

summary(OT_Model11)

## Model 10 and Model 11 Comparison

anova(OT_Model10, OT_Model11)

```

## VAS Analysis

Since we found significant group differences for some acoustic measures between the ALS/PD and Ataxic/HD groups, we continued the hierarchical regression approach from VAS Model 5. VAS Model 6 fit significantly better than VAS Model 5. However, adding in the interactions between Incoord and the acoustic measures did not significantly improve model fit.

### Model 6

```{r}

## Specifying Model 6

VAS_Model6 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + Incoord, data = AcousticData)

## Model 6 Assumption Check

performance::check_model(VAS_Model6)

## Model 6 Summary

summary(VAS_Model6)

## Model 5 and Model 6 Comparison

anova(VAS_Model5, VAS_Model6)


```

### Model 7

```{r}

## Specifying Model 7

VAS_Model7 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25, data = AcousticData)

## Model 7 Assumption Check

performance::check_model(VAS_Model7)

## Model 7 Summary

summary(VAS_Model7)

## Model 6 and Model 7 Comparison

anova(VAS_Model6, VAS_Model7)


```

### Model 8

```{r}

## Specifying Model 8

VAS_Model8 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75, data = AcousticData)

## Model 8 Assumption Check

performance::check_model(VAS_Model8)

## Model 8 Summary

summary(VAS_Model8)

## Model 7 and Model 8 Comparison

anova(VAS_Model7, VAS_Model8)


```

### Model 9

```{r}

## Specifying Model 9

VAS_Model9 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b, data = AcousticData)

## Model 9 Assumption Check

performance::check_model(VAS_Model9)

## Model 9 Summary

summary(VAS_Model9)

## Model 8 and Model 9 Comparison

anova(VAS_Model8, VAS_Model9)

```

### Model 10

```{r}

## Specifying Model 10

VAS_Model10 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b, data = AcousticData)

## Model 10 Assumption Check

performance::check_model(VAS_Model10)

## Model 10 Summary

summary(VAS_Model10)

## Model 9 and Model 10 Comparison

anova(VAS_Model9, VAS_Model10)

```
### Model 11

```{r}

## Specifying Model 11

VAS_Model11 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b + Incoord*vowel_ED_b, data = AcousticData)

## Model 11 Assumption Check

performance::check_model(VAS_Model11)

## Model 11 Summary

summary(VAS_Model11)

## Model 10 and Model 11 Comparison

anova(VAS_Model10, VAS_Model11)

```

### New Final VAS Model

Since VAS Model 6 was significantly better fit than Model 5, we fit a new final parsimonious model for VAS to the data (VAS ~ VSA_b + Incoord) and compared that to the old final VAS model (VAS ~ VSA_b). The new final model was not a significantly better fit than the old final model. Thus the old final model (VAS ~ VSA_b) is retained.

```{r}

## Specifying New Final VAS Model

VAS_Model_newfinal <- lm(VAS ~ VSA_b + Incoord, data = AcousticData)

## New Final VAS Model Assumption Check

performance::check_model(VAS_Model_newfinal)

## New Final VAS Model Summary

summary(VAS_Model_newfinal)

## Comparison to Old Final Model

anova(VAS_Model_final, VAS_Model_newfinal)

```

## Testing Individual Predictors

Because some correlations between predictors are high. We run a series of simple linear regressions to test if the predictors significantly predict each intelligibility measure on their own (VSD 25 and VSA models were already complete in our initial model comparison approach)

### Orthographic Transcription Models

```{r}

# OT ~ VSD 75

OT_vsd75_model <- lm(transAcc ~ Hull_bVSD_75, data = AcousticData)
performance::check_model(OT_vsd75_model)
summary(OT_vsd75_model)

# OT ~ Hull

OT_hull_model <- lm(transAcc ~ Hull_b, data = AcousticData)
performance::check_model(OT_hull_model)
summary(OT_hull_model)

# OT ~ Corner Dispersion

OT_disp_model <- lm(transAcc ~ vowel_ED_b, data = AcousticData)
performance::check_model(OT_disp_model)
summary(OT_disp_model)

```

### VAS Models

```{r}

# VAS ~ VSD 75

VAS_vsd75_model <- lm(VAS ~ Hull_bVSD_75, data = AcousticData)
performance::check_model(VAS_vsd75_model)
summary(VAS_vsd75_model)

# VAS ~ Hull

VAS_hull_model <- lm(VAS ~ Hull_b, data = AcousticData)
performance::check_model(VAS_hull_model)
summary(VAS_hull_model)

# VAS ~ Corner Dispersion

VAS_disp_model <- lm(VAS ~ vowel_ED_b, data = AcousticData)
performance::check_model(VAS_disp_model)
summary(VAS_disp_model)


```
# Listener Demographic Information

```{r}

ListenerDemo <- Listeners %>%
  furniture::table1(age, gender, race, ethnicity)

ListenerDemo

```

# Speaker Demographics

```{r}

SpeakerDemo <- AcousticData %>%
  dplyr::select(c(Speaker, Sex, Etiology))

Ages <- rio::import("Prepped Data/Speaker Ages.xlsx")

SpeakerDemo <- full_join(SpeakerDemo, Ages, by = "Speaker")

SpeakerDemoInfo <- SpeakerDemo %>%
  furniture::table1(Sex, Etiology, Age, na.rm = F)

SpeakerDemoInfo

SpeakerDemo %>%
  dplyr::summarize(mean_age = mean(Age, na.rm = T), age_sd = sd(Age, na.rm = T), age_range = range(Age, na.rm = T))

```

