---
title: "Motor Speech Conference"
output: html_notebook
---

This is the code for the statistical analysis for Motor Speech Conference 2022.

# Loading Packages

```{r}

library(rio)
library(tidyverse)
library(irr)
library(performance)
library(car)
library(ggpubr)
library(corrr)
library(ggridges)
library(furniture)

```

# Upload Datasets

```{r}

Reliability <- rio::import("Prepped Data/Reliability Data.csv")
AcousticData <- rio::import("Prepped Data/AcousticMeasures.csv") 

AcousticData <- AcousticData %>%
  dplyr::filter(!grepl("_rel", Speaker)) %>%
  dplyr::select(c(Speaker, Sex, Etiology, vowel_ED_b, VSA_b, Hull_b, Hull_bVSD_25, Hull_bVSD_50, Hull_bVSD_75, VAS, transAcc))

Listeners <- rio::import("Prepped Data/Listener_Demographics.csv") %>%
  dplyr::select(!c(StartDate:proloficID, Q2.4_6_TEXT, Q3.2_8_TEXT, AudioCheck:EP3))

Listeners$race[Listeners$Q3.3_7_TEXT == "Native American/ African amercing"] <- "Biracial or Multiracial"
```


# Inter-rater Reliability

Two raters (the first two authors) completed vowel segmentation for the speakers. To calculate inter-rater reliability, 20% of the speakers were segmented again by the other rater. Two-way intraclass coefficients were computed for the extracted F1 and F2 from the temporal midpoint of the vowel segments. Since only one set of ratings will be used in the data analysis, we focus on the single ICC results and interpretation. However, we also report the average ICC values to be comprehensive.

```{r}

## Creating new data frames to calculate ICC for extracted F1 and F2 values

F1_Rel <- Reliability %>%
  dplyr::select(c(F1, F1_rel))

F2_Rel <- Reliability %>%
  dplyr::select(c(F2, F2_rel))
  
## Single ICC for F1
Single_F1 <- irr::icc(F1_Rel, model = "twoway", type = "agreement", unit = "single")

## Average ICC for F1
Average_F1 <- irr::icc(F1_Rel, model = "twoway", type = "agreement", unit = "average")

## Single ICC for F2
Single_F2 <- irr::icc(F2_Rel, model = "twoway", type = "agreement", unit = "single")

## Average ICC for F2
Average_F2 <- irr::icc(F2_Rel, model = "twoway", type = "agreement", unit = "average")

## Inter-rater reliability results and interpretation

print(paste("Single ICC for F1 is ", round(Single_F1$value, digits = 3), ". ", 
            "The 95% CI is [", round(Single_F1$lbound, digits = 3), " - ", round(Single_F1$ubound, digits = 3), "].", sep = ""))

print(paste("Single ICC for F2 is ", round(Single_F2$value, digits = 3), ". ", 
            "The 95% CI is [", round(Single_F2$lbound, digits = 3), " - ", round(Single_F2$ubound, digits = 3), "].", sep = ""))

print(paste("Average ICC for F1 is ", round(Average_F1$value, digits = 3), ". ", 
            "The 95% CI is [", round(Average_F1$lbound, digits = 3), " - ", round(Average_F1$ubound, digits = 3), "].", sep = ""))

print(paste("Average ICC for F2 is ", round(Average_F2$value, digits = 3), ". ", 
            "The 95% CI is [", round(Average_F2$lbound, digits = 3), " - ", round(Average_F2$ubound, digits = 3), "].", sep = ""))

print("Thus, interrater reliability for the extracted F1 and F2 values from the vowel segments was good to excellent.")

## Removing extra data frames from environment

rm(F1_Rel, F2_Rel, Reliability, Single_F1, Single_F2, Average_F1, Average_F2)

```


# Descriptive Statistics

## Means and SD

```{r}

Descriptives <- AcousticData %>%
  dplyr::group_by(Sex, Etiology) %>%
  dplyr::summarize(VSA_mean = mean(VSA_b, na.rm =T), VSA_sd = sd(VSA_b, na.rm = T),
                   Disp_mean = mean(vowel_ED_b, na.rm =T), Disp_sd = sd(vowel_ED_b, na.rm =T),
                   Hull_mean = mean(Hull_b, na.rm =T), Hull_sd = sd(Hull_b, na.rm =T),
                   VSD25_mean = mean(Hull_bVSD_25, na.rm =T), VSD25_sd = sd(Hull_bVSD_25, na.rm =T),
                   VSD50_mean = mean(Hull_bVSD_50, na.rm =T), VSD50_sd = sd(Hull_bVSD_50, na.rm =T),
                   VSD75_mean = mean(Hull_bVSD_75, na.rm =T), VSD75_sd = sd(Hull_bVSD_75, na.rm =T),
                   VAS_mean = mean(VAS, na.rm =T), VAS_sd = sd(VAS, na.rm =T),
                   OT_mean = mean(transAcc, na.rm =T), OT_sd = sd(transAcc, na.rm =T))

DescriptivesbySex <- AcousticData %>%
  dplyr::group_by(Sex) %>%
  dplyr::summarize(VSA_mean = mean(VSA_b, na.rm =T), VSA_sd = sd(VSA_b, na.rm = T),
                   Disp_mean = mean(vowel_ED_b, na.rm =T), Disp_sd = sd(vowel_ED_b, na.rm =T),
                   Hull_mean = mean(Hull_b, na.rm =T), Hull_sd = sd(Hull_b, na.rm =T),
                   VSD25_mean = mean(Hull_bVSD_25, na.rm =T), VSD25_sd = sd(Hull_bVSD_25, na.rm =T),
                   VSD50_mean = mean(Hull_bVSD_50, na.rm =T), VSD50_sd = sd(Hull_bVSD_50, na.rm =T),
                   VSD75_mean = mean(Hull_bVSD_75, na.rm =T), VSD75_sd = sd(Hull_bVSD_75, na.rm =T),
                   VAS_mean = mean(VAS, na.rm =T), VAS_sd = sd(VAS, na.rm =T),
                   OT_mean = mean(transAcc, na.rm =T), OT_sd = sd(transAcc, na.rm =T))

DescriptivesbyEtiology <- AcousticData %>%
  dplyr::group_by(Etiology) %>%
  dplyr::summarize(VSA_mean = mean(VSA_b, na.rm =T), VSA_sd = sd(VSA_b, na.rm = T),
                   Disp_mean = mean(vowel_ED_b, na.rm =T), Disp_sd = sd(vowel_ED_b, na.rm =T),
                   Hull_mean = mean(Hull_b, na.rm =T), Hull_sd = sd(Hull_b, na.rm =T),
                   VSD25_mean = mean(Hull_bVSD_25, na.rm =T), VSD25_sd = sd(Hull_bVSD_25, na.rm =T),
                   VSD50_mean = mean(Hull_bVSD_50, na.rm =T), VSD50_sd = sd(Hull_bVSD_50, na.rm =T),
                   VSD75_mean = mean(Hull_bVSD_75, na.rm =T), VSD75_sd = sd(Hull_bVSD_75, na.rm =T),
                   VAS_mean = mean(VAS, na.rm =T), VAS_sd = sd(VAS, na.rm =T),
                   OT_mean = mean(transAcc, na.rm =T), OT_sd = sd(transAcc, na.rm =T))

Descriptives

DescriptivesbySex

DescriptivesbyEtiology

```

## Correlations among variables

```{r}

CorrMatrix1 <- AcousticData %>%
  dplyr::select(VSA_b, vowel_ED_b, Hull_b, Hull_bVSD_25, Hull_bVSD_50, Hull_bVSD_75, VAS, transAcc) %>%
  corrr::correlate()

CorrMatrix1

```


## Data Vis

### Group Comparisons: Ridgeline Plots

```{r}

# Vowel Space Area Distribution by Etiology

VSA_plot <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Area (Bark)") +
  theme_classic()

## Corner Dispersion Distribution by Etiology

disp_plot <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Corner Dispersion (Bark)") +
  theme_classic()

# Vowel Space Hull Distribution by Etiology

Hull_plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Hull (Bark)") +
  theme_classic()

# Vowel Space Density 25 Distribution by Etiology

vsd_25_plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Denisty 25 (Bark)") +
  theme_classic()

# Vowel Space Density 50 Distribution by Etiology

vsd_50_plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_50,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Denisty 50 (Bark)") +
  theme_classic()

# Vowel Space Density 75 Distribution by Etiology

vsd_75_plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  expand_limits(x=0) +
  xlab("Vowel Space Denisty 75 (Bark)") +
  theme_classic()

# Visual Analog Scale Intelligibility Rating Distribution by Etiology

VAS_plot <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Visual Analog Scale (VAS)") +
  theme_classic()

# Orthographic Transcription Score Distribution by Etiology

OT_plot <- AcousticData %>%
  ggplot() +
  aes(x = transAcc,
      y = Etiology,
      color = Etiology,
      fill = Etiology) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  coord_cartesian(xlim = c(0, 100)) +
  xlab("Orthographic Transcription Scores") +
  theme_classic()

# Creating Distributions Figure

Distributions <- ggpubr::ggarrange(VSA_plot, disp_plot, Hull_plot, vsd_25_plot, vsd_50_plot, vsd_75_plot, VAS_plot, OT_plot,
                  ncol = 2,
                  nrow = 4)

# Saving Distribution Figure

ggsave("Plots/Distribution.png", plot = last_plot(), width = 10, height = 10, unit = "in")


```

### Scatter Plots

```{r}

# OT Scatterplots

OT_VSA <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#E69F00", fill = "light grey") +
  xlab(expression("Vowel Space Area (Bark"^2*")")) +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

OT_disp <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#56B4E9", fill = "light grey") +
  xlab("Corner Dispersion (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_Hull <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#009E73", fill = "light grey") +
  xlab(expression("Vowel Space Hull (Bark"^2*")")) +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd25 <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#0072B2", fill = "light grey") +
  xlab(expression("Vowel Space Density"[25]*" (Bark"^2*")")) +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd75 <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#D55E00", fill = "light grey") +
  xlab(expression("Vowel Space Density"[75]*" (Bark"^2*")")) +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

# Creating OT Scatterplot Figure

OT_Scatterplots <- ggpubr::ggarrange(OT_VSA, OT_disp, OT_Hull, OT_vsd25, OT_vsd75,
                                     ncol = 3, nrow = 2)

## Saving OT Scatterplot Figure

ggsave("Plots/OT_Scatterplots.png", plot = last_plot(), width = 15, height = 15, unit = "in", scale = 0.6)

# VAS Scatterplots

VAS_VSA <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = VAS) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#E69F00", fill = "light grey") +
  xlab(expression("Vowel Space Area (Bark"^2*")")) +
  ylab("VAS Rating") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

VAS_disp <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = VAS) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#56B4E9", fill = "light grey") +
  xlab("Corner Dispersion (Bark)") +
  ylab("VAS Rating") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_Hull <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = VAS) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#009E73", fill = "light grey") +
  xlab(expression("Vowel Space Hull (Bark"^2*")")) +
  ylab("VAS Rating") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd25 <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = VAS) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#0072B2", fill = "light grey") +
  xlab(expression("Vowel Space Density"[25]*" (Bark"^2*")")) +
  ylab("VAS Rating") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd75 <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = VAS) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#D55E00", fill = "light grey") +
  xlab(expression("Vowel Space Density"[75]*" (Bark"^2*")")) +
  ylab("VAS Rating") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

# Creating VAS Scatterplot Figure

VAS_Scatterplots <- ggpubr::ggarrange(VAS_VSA, VAS_disp, VAS_Hull, VAS_vsd25, VAS_vsd75,
                  ncol = 3, nrow = 2)

## Saving VAS Scatterplot Figure

ggsave("Plots/VAS_Scatterplots.png", plot = last_plot(), width = 15, height = 15, unit = "in", scale = .6)

```


### OT ~ VAS Plot

```{r}

# VAS ~ OT Plot

VAS_OT <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = transAcc) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#CC79A7", fill = "light grey") + 
  ggtitle("Relationship Between Intelligibility Measures") +
  xlab("VAS Rating") +
  ylab("Percent Words Correct") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Saving OT ~ VAS Plot

ggsave("Plots/OTvas.png", plot = last_plot(), width = 8, height = 8, unit = "in", scale = 0.6)
  
# OT ~ VAS by Sex

VAS_OT_Sex <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = transAcc,
      color = Sex) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("VAS Rating") +
  ylab("Percent Words Correct") +
  theme_classic()

ggsave("Plots/OTvas_sex.png", plot = last_plot(), width = 5, height = 5, unit = "in", scale = 0.6)

# OT ~ VAS by Etiology

VAS_OT_Etiology <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = transAcc,
      color = Etiology) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("VAS Rating") +
  ylab("Percent Words Correct") +
  theme_classic()

ggsave("Plots/OTvas_etiology.png", plot = last_plot(), width = 5, height = 5, unit = "in", scale = 0.6)

```


# Group Comparisons

## VSA

```{r}

## Specify the Model
VSA_group <- aov(VSA_b ~ Etiology, data = AcousticData)

## Assumption Check

plot(VSA_group, 1)
plot(VSA_group, 2)
car::leveneTest(VSA_group)
VSA_residuals <- residuals(object = VSA_group)
shapiro.test(VSA_residuals)

## Model Results

summary(VSA_group)

## Kruskal-Wallis Test 

kruskal.test(VSA_b ~ Etiology, data = AcousticData)

## Pairwise Comparisons

pairwise.wilcox.test(AcousticData$VSA_b, AcousticData$Etiology, p.adjust.method = "bonferroni")

```

## Corner Dispersion

```{r}

## Specify the Model
disp_group <- aov(vowel_ED_b ~ Etiology, data = AcousticData)

## Assumption Check

plot(disp_group, 1)
plot(disp_group, 2)
car::leveneTest(disp_group)
disp_residuals <- residuals(object = disp_group)
shapiro.test(disp_residuals)

## Model Results

summary(disp_group)

```

## Hull

```{r}

## Specify the Model
hull_group <- aov(Hull_b ~ Etiology, data = AcousticData)

## Assumption Check

plot(hull_group, 1)
plot(hull_group, 2)
car::leveneTest(hull_group)
hull_residuals <- residuals(object = hull_group)
shapiro.test(hull_residuals)

## Model Results

summary(hull_group)

```

## VSD 25

```{r}

## Specify the Model
vsd25_group <- aov(Hull_bVSD_25 ~ Etiology, data = AcousticData)

## Assumption Check

plot(vsd25_group, 1)
plot(vsd25_group, 2)
car::leveneTest(vsd25_group)
vsd25_residuals <- residuals(object = vsd25_group)
shapiro.test(vsd25_residuals)

## Model Summary

summary(vsd25_group)


```


## VSD 50

```{r}

## Specify the Model
vsd50_group <- aov(Hull_bVSD_50 ~ Etiology, data = AcousticData)

## Assumption Check

plot(vsd50_group, 1)
plot(vsd50_group, 2)
car::leveneTest(vsd50_group)
vsd50_residuals <- residuals(object = vsd50_group)
shapiro.test(vsd50_residuals)

## Model Results

summary(vsd50_group)

## Kruskal Wallis Test

kruskal.test(Hull_bVSD_50 ~ Etiology, data = AcousticData)

```

## VSD 75

```{r}

## Specify the Model
vsd75_group <- aov(Hull_bVSD_75 ~ Etiology, data = AcousticData)

## Assumption Check

plot(vsd75_group, 1)
plot(vsd75_group, 2)
car::leveneTest(vsd75_group)
vsd75_residuals <- residuals(object = vsd75_group)
shapiro.test(vsd75_residuals)

## Model Summary

summary(vsd75_group)

## Kruskal Wallis

kruskal.test(Hull_bVSD_75 ~ Etiology, data = AcousticData)

```

## VAS 

```{r}

## Specify the Model
VAS_group <- aov(VAS ~ Sex*Etiology, data = AcousticData)

## Assumption Check

plot(VAS_group, 1)
plot(VAS_group, 2)
car::leveneTest(VAS_group)
VAS_residuals <- residuals(object = VAS_group)
shapiro.test(VAS_residuals)

## Model Results

summary(VAS_group)

```

## OT

```{r}

## Specify the Model
OT_group <- aov(transAcc ~ Sex*Etiology, data = AcousticData)

## Assumption Check

plot(OT_group, 1)
plot(OT_group, 2)
car::leveneTest(OT_group)
OT_residuals <- residuals(object = OT_group)
shapiro.test(OT_residuals)

## Model Results

summary(OT_group)

```

# Modeling Intelligibility

## Orthographic Transcriptions

### Model 1

```{r}

# Specifying Model 1

OT_Model1 <- lm(transAcc ~ Hull_bVSD_25, data = AcousticData)

## Model 1 Assumptions 

performance::check_model(OT_Model1)

## Model 1 Summary

summary(OT_Model1)

```

### Model 2

```{r}

## Specifying Model 2

OT_Model2 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75, data = AcousticData)

## Model 2 Assumption Check

performance::check_model(OT_Model2)

## Model 2 Summary

summary(OT_Model2)

## Model 1 and Model 2 Comparison

anova(OT_Model1, OT_Model2)

```

### Model 3

```{r}

## Specifying Model 3

OT_Model3 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b, data = AcousticData)

## Model 3 Assumption Check

performance::check_model(OT_Model3)

## Model 3 Summary

summary(OT_Model3)

## Model 2 and Model 3 Comparison

anova(OT_Model2, OT_Model3)

```

### Model 4

```{r}

## Specifying Model 4

OT_Model4 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b, data = AcousticData)

## Model 4 Assumption Check

performance::check_model(OT_Model4)

## Model 4 Summary

summary(OT_Model4)

## Model 3 and Model 4 Comparison

anova(OT_Model3, OT_Model4)

```
### Model 5

```{r}

## Specifying Model 5

OT_Model5 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b, data = AcousticData)

## Model 4 Assumption Check

performance::check_model(OT_Model5)

## Model 4 Summary

summary(OT_Model5)

## Model 3 and Model 4 Comparison

anova(OT_Model4, OT_Model5)

```


### Final Model

```{r}

## Specifying Final Model

OT_Model_final <- lm(transAcc ~ VSA_b, data = AcousticData)

## Final Model Assumption Check

performance::check_model(OT_Model_final)

## Final Model Summary

summary(OT_Model_final)

```


## VAS Models

### Model 1

```{r}

# Specifying Model 1

VAS_Model1 <- lm(VAS ~ Hull_bVSD_25, data = AcousticData)

## Model 1 Assumptions 

performance::check_model(VAS_Model1)

## Model 1 Summary

summary(VAS_Model1)

```

### Model 2

```{r}

## Specifying Model 2

VAS_Model2 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75, data = AcousticData)

## Model 2 Assumption Check

performance::check_model(VAS_Model2)

## Model 2 Summary

summary(VAS_Model2)

## Model 1 and Model 2 Comparison

anova(VAS_Model1, VAS_Model2)

```

### Model 3

```{r}

## Specifying Model 3

VAS_Model3 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b, data = AcousticData)

## Model 3 Assumption Check

performance::check_model(VAS_Model3)

## Model 3 Summary

summary(VAS_Model3)

## Model 2 and Model 3 Comparison

anova(VAS_Model2, VAS_Model3)

```

### Model 4

```{r}

## Specifying Model 4

VAS_Model4 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b, data = AcousticData)

## Model 4 Assumption Check

performance::check_model(VAS_Model4)

## Model 4 Summary

summary(VAS_Model4)

## Model 3 and Model 4 Comparison

anova(VAS_Model3, VAS_Model4)

```

### Model 5

```{r}

## Specifying Model 5

VAS_Model5 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b, data = AcousticData)

## Model 5 Assumption Check

performance::check_model(VAS_Model5)

## Model 5 Summary

summary(VAS_Model5)

## Model 4 and Model 5 Comparison

anova(VAS_Model4, VAS_Model5)

```


### Final Model

```{r}

## Specifying Final Model

VAS_Model_final <- lm(VAS ~ VSA_b, data = AcousticData)

## Final Model Assumption Check

performance::check_model(VAS_Model_final)

## Final Model Summary

summary(VAS_Model_final)

```
## Relationship between OT and VAS

### Tested Model

```{r}

# Specify Model

OT_VAS_model <- lm(transAcc ~ VAS*Etiology + VAS*Sex, data = AcousticData)

# Assumption Check

performance::check_model(OT_VAS_model)

# Model Results

summary(OT_VAS_model)

```

## Final Model

```{r}

# Specify Final Model

OT_VAS_final <- lm(transAcc ~ VAS, data = AcousticData)

# Model Results

summary(OT_VAS_final)

```

# Alternate Analysis

In this alternate analysis, we are looking at the relationship between these acoustic measures with speech intelligibility for the ALS/PD and the HD/Ataxic speakers separately. We create a new variable called Incoord, where the ALS/PD Speakers are set as the reference group (in order to compare to the Ataxic/HD Speaker Group). Group Comparisons, additional data visualizations, and further linear model comparisons are completed.

## Data Prep

```{r}

# Creating a new variable in AcousticData. Incoord is a dummy variable. ALS/PD Speakers = 0, HD/Ataxic = 1

AcousticData <- AcousticData %>%
  dplyr::mutate(Incoord = case_when(Etiology == "HD" ~ 1,
                                    Etiology == "Ataxic" ~ 1,
                                    TRUE ~ 0)) %>%
  dplyr::mutate(Incoord = as.factor(Incoord))

```

## Descriptives 

```{r}

Descriptives_ALS.PD_incoord <- AcousticData %>%
  dplyr::group_by(Incoord) %>%
  dplyr::summarize(VSA_mean = mean(VSA_b, na.rm =T), VSA_sd = sd(VSA_b, na.rm = T),
                   Disp_mean = mean(vowel_ED_b, na.rm =T), Disp_sd = sd(vowel_ED_b, na.rm =T),
                   Hull_mean = mean(Hull_b, na.rm =T), Hull_sd = sd(Hull_b, na.rm =T),
                   VSD25_mean = mean(Hull_bVSD_25, na.rm =T), VSD25_sd = sd(Hull_bVSD_25, na.rm =T),
                   VSD50_mean = mean(Hull_bVSD_50, na.rm =T), VSD50_sd = sd(Hull_bVSD_50, na.rm =T),
                   VSD75_mean = mean(Hull_bVSD_75, na.rm =T), VSD75_sd = sd(Hull_bVSD_75, na.rm =T),
                   VAS_mean = mean(VAS, na.rm =T), VAS_sd = sd(VAS, na.rm =T),
                   OT_mean = mean(transAcc, na.rm =T), OT_sd = sd(transAcc, na.rm =T))

Descriptives_ALS.PD_incoord

```

## Data Vis

### Group Comparison Ridgeline Plots

```{r}

# VSA Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

VSA_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Area (Bark)") +
  theme_classic()

# Corner Dispersion Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

disp_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Corner Dispersion (Bark)") +
  theme_classic()

# Hull Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

hull_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Hull (Bark)") +
  theme_classic()

# VSD 25 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd25_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 25 (Bark)") +
  theme_classic()

# VSD 50 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd50_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_50,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 50 (Bark)") +
  theme_classic()

# VSD 75 Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

vsd75_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Vowel Space Desnity 75 (Bark)") +
  theme_classic()

# Orthographic Transcription Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

OT_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = transAcc,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Percent Words Correct") +
  theme_classic()

# Visual Analog Scale Intelligibility Rating Ridgeline Plot by Incoordination (Incoord = 0 are ALS/PD Speakers, Incoord == 1 are HD/Ataxic Speakers)

VAS_incoord.plot <- AcousticData %>%
  ggplot() +
  aes(x = VAS,
      y = Incoord,
      color = Incoord,
      fill = Incoord) +
  geom_density_ridges(jittered_points = T, 
                      position = position_points_jitter(width = 0.01, height = 0), 
                      point_shape = '|', 
                      point_size = 5,
                      point_alpha = 1,
                      alpha = 0.7,
                      scale = .7) +
  xlab("Ratigns") +
  theme_classic()

# Creating Distributions by Incoord Figure

Distributions_incoord <- ggpubr::ggarrange(VSA_incoord.plot, disp_incoord.plot, hull_incoord.plot, vsd25_incoord.plot, vsd50_incoord.plot, vsd75_incoord.plot, OT_incoord.plot, VAS_incoord.plot,
                  ncol = 2,
                  nrow = 4)

# Saving Distribution Figure

ggsave("Plots/Distribution_incoord.png", plot = last_plot(), width = 10, height = 10, unit = "in")

```

### Scatterplots

```{r}

# OT Scatterplots by Incoord

OT_VSA.incoord <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Area (Bark)") +
  ylab("Percent Words Correct") +
  ggtitle("Orthographic Transcription") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

OT_disp.incoord <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Corner Dispersion (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_Hull.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Hull (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd25.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 25 (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

OT_vsd75.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = transAcc,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 75 (Bark)") +
  ylab("Percent Words Correct") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()


# VAS Scatterplots

VAS_VSA.incoord <- AcousticData %>%
  ggplot() +
  aes(x = VSA_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Area (Bark)") +
  ylab("VAS Score") +
  ggtitle("Visual Analog Scale") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

VAS_disp.incoord <- AcousticData %>%
  ggplot() +
  aes(x = vowel_ED_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Corner Dispersion (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_Hull.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_b,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Hull (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd25.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_25,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 25 (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()

VAS_vsd75.incoord <- AcousticData %>%
  ggplot() +
  aes(x = Hull_bVSD_75,
      y = VAS,
      color = Incoord) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  xlab("Vowel Space Density 75 (Bark)") +
  ylab("VAS Score") +
  coord_cartesian(ylim = c(0,100)) +
  theme_classic()


# Creating Scatterplot Figure

Scatterplots_incoord <- ggpubr::ggarrange(OT_VSA.incoord, VAS_VSA.incoord, OT_disp.incoord, VAS_disp.incoord, OT_Hull.incoord, VAS_Hull.incoord, OT_vsd25.incoord, VAS_vsd25.incoord, OT_vsd75.incoord, VAS_vsd75.incoord,
                  ncol = 2,
                  nrow = 5)

## Saving Scatterplot Figure

ggsave("Plots/Scatterplots_incoord.png", plot = last_plot(), width = 15, height = 20, unit = "in", scale = .5)


```


## Group Comparisons

### Two Dataframes for Incoord Groups

```{r}

coord.group <- AcousticData %>%
  dplyr::filter(Incoord == 0)

incoord.group <- AcousticData %>%
  dplyr::filter(Incoord == 1)

```


### VSA

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(VSA_b[Incoord == 0]))
with(AcousticData, shapiro.test(VSA_b[Incoord == 1]))

## Equal Variance Check

res.ftest.VSA <- var.test(VSA_b ~ Incoord, data = AcousticData)
res.ftest.VSA

# Model Results

VSA_b_t <- t.test(incoord.group$VSA_b, coord.group$VSA_b, var.equal = T)
VSA_b_t

```

### Corner Dispersion

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(vowel_ED_b[Incoord == 0]))
with(AcousticData, shapiro.test(vowel_ED_b[Incoord == 1]))

## Equal Variance Check

res.ftest.disp <- var.test(vowel_ED_b ~ Incoord, data = AcousticData)
res.ftest.disp

# Model Results

disp_t <- t.test(incoord.group$vowel_ED_b, coord.group$vowel_ED_b, var.equal = T)
disp_t

```

### Hull

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_b[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_b[Incoord == 1]))

## Equal Variance Check

res.ftest.hull <- var.test(Hull_b ~ Incoord, data = AcousticData)
res.ftest.hull

# Model Results

hull_t <- t.test(incoord.group$Hull_b, coord.group$Hull_b, var.equal = T)
hull_t

```

### VSD 25

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_25[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_25[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd25 <- var.test(Hull_bVSD_25 ~ Incoord, data = AcousticData)
res.ftest.vsd25

# Model Results

vsd25_t <- t.test(incoord.group$Hull_bVSD_25, coord.group$Hull_bVSD_25, var.equal = T)
vsd25_t

```

### VSD 50

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_50[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_50[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd50 <- var.test(Hull_bVSD_50 ~ Incoord, data = AcousticData)
res.ftest.vsd50

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

vsd50_MW <- wilcox.test(Hull_bVSD_50 ~ Incoord, data = AcousticData)
vsd50_MW

```

### VSD 75

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(Hull_bVSD_75[Incoord == 0]))
with(AcousticData, shapiro.test(Hull_bVSD_75[Incoord == 1]))

## Equal Variance Check

res.ftest.vsd75 <- var.test(Hull_bVSD_75 ~ Incoord, data = AcousticData)
res.ftest.vsd75

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

vsd75_MW <- wilcox.test(Hull_bVSD_75 ~ Incoord, data = AcousticData)
vsd75_MW


```

### Orthographic Transcription Scores

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(transAcc[Incoord == 0]))
with(AcousticData, shapiro.test(transAcc[Incoord == 1]))

## Equal Variance Check

res.ftest.OT <- var.test(transAcc ~ Incoord, data = AcousticData)
res.ftest.OT

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

OT_MW <- wilcox.test(transAcc ~ Incoord, data = AcousticData)
OT_MW

```

### VAS

```{r}

# Assumption Check

## Checking Normality

with(AcousticData, shapiro.test(VAS[Incoord == 0]))
with(AcousticData, shapiro.test(VAS[Incoord == 1]))

## Equal Variance Check

res.ftest.VAS <- var.test(VAS ~ Incoord, data = AcousticData)
res.ftest.VAS

# Model Results (Mann-Whitney U test conducted since assumption of normality is violated)

OT_MW <- wilcox.test(VAS ~ Incoord, data = AcousticData)
OT_MW

```

## OT Analysis

Since we found significant group differences for some acoustic measures between the ALS/PD and Ataxic/HD groups, we continued the heirarichal regression approach from OT Model 5. Adding in the Incoord predictor along with the interactions between the acoustic measures did not significantly improve model fit. So our original final OT model is retained.

### Model 6

```{r}

## Specifying Model 6

OT_Model6 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + Incoord, data = AcousticData)

## Model 6 Assumption Check

performance::check_model(OT_Model6)

## Model 6 Summary

summary(OT_Model6)

## Model 5 and Model 6 Comparison

anova(OT_Model5, OT_Model6)

```
### Model 7

```{r}

## Specifying Model 7

OT_Model7 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25, data = AcousticData)

## Model 7 Assumption Check

performance::check_model(OT_Model7)

## Model 7 Summary

summary(OT_Model7)

## Model 6 and Model 7 Comparison

anova(OT_Model6, OT_Model7)

```

### Model 8

```{r}

## Specifying Model 8

OT_Model8 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75, data = AcousticData)

## Model 8 Assumption Check

performance::check_model(OT_Model8)

## Model 8 Summary

summary(OT_Model8)

## Model 7 and Model 8 Comparison

anova(OT_Model7, OT_Model8)

```

### Model 9

```{r}

## Specifying Model 9

OT_Model9 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b, data = AcousticData)

## Model 9 Assumption Check

performance::check_model(OT_Model9)

## Model 9 Summary

summary(OT_Model9)

## Model 8 and Model 9 Comparison

anova(OT_Model8, OT_Model9)

```

### Model 10

```{r}

## Specifying Model 10

OT_Model10 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b, data = AcousticData)

## Model 10 Assumption Check

performance::check_model(OT_Model10)

## Model 10 Summary

summary(OT_Model10)

## Model 9 and Model 10 Comparison

anova(OT_Model9, OT_Model10)

```
### Model 11

```{r}

## Specifying Model 11

OT_Model11 <- lm(transAcc ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b + Incoord*vowel_ED_b, data = AcousticData)

## Model 11 Assumption Check

performance::check_model(OT_Model11)

## Model 11 Summary

summary(OT_Model11)

## Model 10 and Model 11 Comparison

anova(OT_Model10, OT_Model11)

```

## VAS Analysis

Since we found significant group differences for some acoustic measures between the ALS/PD and Ataxic/HD groups, we continued the hierarchical regression approach from VAS Model 5. VAS Model 6 fit significantly better than VAS Model 5. However, adding in the interactions between Incoord and the acoustic measures did not significantly improve model fit.

### Model 6

```{r}

## Specifying Model 6

VAS_Model6 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + Incoord, data = AcousticData)

## Model 6 Assumption Check

performance::check_model(VAS_Model6)

## Model 6 Summary

summary(VAS_Model6)

## Model 5 and Model 6 Comparison

anova(VAS_Model5, VAS_Model6)


```

### Model 7

```{r}

## Specifying Model 7

VAS_Model7 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25, data = AcousticData)

## Model 7 Assumption Check

performance::check_model(VAS_Model7)

## Model 7 Summary

summary(VAS_Model7)

## Model 6 and Model 7 Comparison

anova(VAS_Model6, VAS_Model7)


```

### Model 8

```{r}

## Specifying Model 8

VAS_Model8 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75, data = AcousticData)

## Model 8 Assumption Check

performance::check_model(VAS_Model8)

## Model 8 Summary

summary(VAS_Model8)

## Model 7 and Model 8 Comparison

anova(VAS_Model7, VAS_Model8)


```

### Model 9

```{r}

## Specifying Model 9

VAS_Model9 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b, data = AcousticData)

## Model 9 Assumption Check

performance::check_model(VAS_Model9)

## Model 9 Summary

summary(VAS_Model9)

## Model 8 and Model 9 Comparison

anova(VAS_Model8, VAS_Model9)

```

### Model 10

```{r}

## Specifying Model 10

VAS_Model10 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b, data = AcousticData)

## Model 10 Assumption Check

performance::check_model(VAS_Model10)

## Model 10 Summary

summary(VAS_Model10)

## Model 9 and Model 10 Comparison

anova(VAS_Model9, VAS_Model10)

```
### Model 11

```{r}

## Specifying Model 11

VAS_Model11 <- lm(VAS ~ Hull_bVSD_25 + Hull_bVSD_75 + Hull_b + VSA_b + vowel_ED_b + 
                  Incoord + Incoord*Hull_bVSD_25 + Incoord*Hull_bVSD_75 + Incoord*Hull_b + Incoord*VSA_b + Incoord*vowel_ED_b, data = AcousticData)

## Model 11 Assumption Check

performance::check_model(VAS_Model11)

## Model 11 Summary

summary(VAS_Model11)

## Model 10 and Model 11 Comparison

anova(VAS_Model10, VAS_Model11)

```

### New Final VAS Model

Since VAS Model 6 was significantly better fit than Model 5, we fit a new final parsimonious model for VAS to the data (VAS ~ VSA_b + Incoord) and compared that to the old final VAS model (VAS ~ VSA_b). The new final model was not a significantly better fit than the old final model. Thus the old final model (VAS ~ VSA_b) is retained.

```{r}

## Specifying New Final VAS Model

VAS_Model_newfinal <- lm(VAS ~ VSA_b + Incoord, data = AcousticData)

## New Final VAS Model Assumption Check

performance::check_model(VAS_Model_newfinal)

## New Final VAS Model Summary

summary(VAS_Model_newfinal)

## Comparison to Old Final Model

anova(VAS_Model_final, VAS_Model_newfinal)

```

## Testing Individual Predictors

Because some correlations between predictors are high. We run a series of simple linear regressions to test if the predictors significantly predict each intelligibility measure on their own (VSD 25 and VSA models were already complete in our initial model comparison approach)

### Orthographic Transcription Models

```{r}

# OT ~ VSD 75

OT_vsd75_model <- lm(transAcc ~ Hull_bVSD_75, data = AcousticData)
performance::check_model(OT_vsd75_model)
summary(OT_vsd75_model)

# OT ~ Hull

OT_hull_model <- lm(transAcc ~ Hull_b, data = AcousticData)
performance::check_model(OT_hull_model)
summary(OT_hull_model)

# OT ~ Corner Dispersion

OT_disp_model <- lm(transAcc ~ vowel_ED_b, data = AcousticData)
performance::check_model(OT_disp_model)
summary(OT_disp_model)

```

### VAS Models

```{r}

# VAS ~ VSD 75

VAS_vsd75_model <- lm(VAS ~ Hull_bVSD_75, data = AcousticData)
performance::check_model(VAS_vsd75_model)
summary(VAS_vsd75_model)

# VAS ~ Hull

VAS_hull_model <- lm(VAS ~ Hull_b, data = AcousticData)
performance::check_model(VAS_hull_model)
summary(VAS_hull_model)

# VAS ~ Corner Dispersion

VAS_disp_model <- lm(VAS ~ vowel_ED_b, data = AcousticData)
performance::check_model(VAS_disp_model)
summary(VAS_disp_model)


```
# Listener Demographic Information

```{r}

ListenerDemo <- Listeners %>%
  furniture::table1(age, gender, race, ethnicity)

ListenerDemo

```

# Speaker Demographics

```{r}

SpeakerDemo <- AcousticData %>%
  dplyr::select(c(Speaker, Sex, Etiology))

Ages <- rio::import("Prepped Data/Speaker Ages.xlsx")

SpeakerDemo <- full_join(SpeakerDemo, Ages, by = "Speaker")

SpeakerDemoInfo <- SpeakerDemo %>%
  furniture::table1(Sex, Etiology, Age, na.rm = F)

SpeakerDemoInfo

SpeakerDemo %>%
  dplyr::summarize(mean_age = mean(Age, na.rm = T), age_sd = sd(Age, na.rm = T), age_range = range(Age, na.rm = T))

```

